Binary Image Transition State Search Method

### 2022-03-17 manuscript.tex
Added citations

% Figures
\usepackage{graphicx}

@@

More recently, it is increasingly being recognised that transition states are useful in mesoscale or macroscale systems.
Here, the minimum energy barriers provide important lower bounds to the energy input required for transitions to occur.
This has been used to understand failure in structural engineering applications \cite{Panter2019,Hutchinson2018}, for the development of super liquid-repellent surfaces \cite{Zhang2014,Panter2019b}, and investigating locomotion through complex terrain for robotics \cite{Othayoth2020}.
Moreover, it is becoming desirable to tailor elastic deformation transitions to enable technologies such as advanced deployable structures \cite{Filipov2015,Zhai2018}, mechanical sensors and actuators \cite{Bertoldi2017,Truby2016,Chi2022,Bonfanti2020}, and energy absorbers \cite{Shan2015,Giri2021} to name but a few.

Existing transition state search methods generally fall into two categories, single- and double-ended methods.
Single-ended methods are initialised at a single state and attempt to climb to a nearby saddle point.
Examples include eigenvector following \cite{Cerjan1981}, the dimer method \cite{Heyden2005,Kastner2008}, and climbing image methods \cite{E2007,Ren2013}.
Double-ended methods, such as the string method \cite{E2002,E2007} and doubly-nudged elastic band (DNEB) \cite{Trygubenko2004}, involve initialising a chain of states between two minima and then minimising the total energy of the chain.
In addition to approximating the transition state, these are useful for obtaining an estimate for the full transition pathway.

However, a large range of landscapes prove challenging or impossible to explore via these methods.
One key problem arises from the push towards larger and more complex systems \cite{Trefethen2013,Shalf2020,Alexander2020}, resulting in the need to develop algorithms that are more memory and computationally efficient, and can incorporate advanced optimisation strategies such as on-the-fly adaptive remeshing and coarse-graining.
Current double-ended methods require a chain of tens of states to gain an accurate approximation, and have issues with remeshing due to the different number of degrees of freedom for each state.
Single-ended methods are not well suited for identifying specific pathways and can spend a large amount of time searching for undesired transition states.
Another major challenge in studying complex energy landscapes relates to the presence of locally flat or discontinous regions.
Such as when considering patchy \cite{McMullen2018,Nguemaha2018,Chen2018b} and hard-body \cite{Richard2018,Santra2018} interactions in atomistic systems, or collision constraints for macroscopic objects \cite{Wriggers2006}.
Flat zero-modes in the landscape pose issues for single-ended search methods that rely only upon local information.
Specialist treatment can sometimes be used such as in the case of global rotation and translation \cite{Page1988}, but they are thwarted by local zero-modes.
Meanwhile, current methods cannot typically be applied in the case of discontinuous potentials, or where the gradient is prohibitively expensive to compute, because they require continuous, differentiable optimisation functions.

\begin{figure}[htb]
  \includegraphics{fig/toy2d.pdf}
  \caption{\label{fig:toy2d}
    (a) The trajectories of the two states under the BITSS method (orange line) for a simple 2D potential with two minima (blue dots).
        The transition state is shown by a red dot, and the minimum energy pathway by the dashed line.
    (b) A snapshot of the BITSS minimisation showing the driving forces on each state due to the energy constraint, $\bm{F}_E$, and distance constraint, $\bm{F}_D$, with $E_1 < E_2$ and $\mathrm{d}(\bm{x}_1,\bm{x}_2) < d_i$.
    (c) The final configuration of the BITSS method showing the two states in orange, the transition state in red, and the negative curvature eigenvector, $\bm{\hat{\tau}}$.
  }
\end{figure}

In this work we introduce a new method, the Binary Image Transition State Search (BITSS).
This is a double-ended method that uses two states that converge toward a transition state from either side, therefore combining aspects of the two traditional transition state search categories.
We demonstrate how it addresses each of the above challenges using a range of different applications.

@@

To ensure that the transition state is located successfully, the constraint strengths $\kappa_d$ and $\kappa_e$ are updated as the algorithm proceeds using information from the system.
They are initially calculated at the start of each minimisation, and regularly recalculated throughout (once per 100 iterations is used in this work).
They are set using the following equations such that each term in the gradient of \cref{eq:bitss} is of a similar size,
\begin{gather}
  \kappa_e = \frac {\alpha} {2 E_B},
  \label{eq:ke}
  \\
  \kappa_d = \max \left(
    \frac {\sqrt{|\bm{\nabla} E_1|^2 + |\bm{\nabla} E_2|^2}} {2\sqrt{2} \beta d_i} \; , \;
    \frac{E_B}{\beta d_i^2} \right),
  \label{eq:kd}
\end{gather}
where $\bm{\nabla} E_1$ and $\bm{\nabla} E_2$ are the gradients of the energies two states, and $\alpha$ and $\beta$ are parameters with recommended values of $\alpha = 10$ and $\beta = 0.1$ (see supplementary information \footnote{See Supplemental Material at [URL] for derivations of the constraint coefficients, extra details for the example systems, results for the speed test against DNEB, and a comparison of the discontinuous energy landscape with the standard Lennard-Jones cluster}).
Here, $E_B$ is an estimation for the current energy barrier; evaluated using the difference between the highest energy along a linear interpolation between the two states and the average energy of the two states.

@@

%\section{Speed Comparison}
\begin{figure*}[tb]
  \includegraphics{fig/speedtest.pdf}
  \caption{\label{fig:speedtest}
    The convergence to the transition state as a function of the number of gradient calculations using the BITSS (black line) and string methods for (a) a Lennard-Jones seven-particle cluster, (b) cylindrical shell buckling, and (c) wetting of a chemically-striped surface.
    The string method is repeated with a differing number of images along the string, the number of which is listed in the legend in (c).
    The shown configurations correspond to the two minimum energy states and the transition state, marked by an asterisk.
  }
\end{figure*}
As a result of the BITSS method using only two states it has the potential to be significantly faster than chain-of-states approaches which typically use tens of states.
This is tested here by measuring the speed of convergence to the transition state for BITSS in comparison with the string method and DNEB for several representative examples.

@@

%\section{Adaptive mesh}
\begin{figure}[tb]
  \includegraphics{fig/adaptivemesh.pdf}
  \caption{\label{fig:adaptivemesh}
    (a) Snapshots of the BITSS method applied to the buckling of a cylinder with a changing mesh.
        The radial displacement relative the unbuckled cylinder is shown, as well as the underlying triangular mesh.
    (b) BITSS applied to the striped wetting example with different resolutions for the two states.
        Each grid cell denotes 50x50 lattice nodes.
        The zoomed axes shows the difference in the fluid interface between the two final states, as well as the approximated transition state (solid black line).
        This is compared to the transition state found using a high resolution (dashed line).
  }
\end{figure}

@@

%\section{Zero-curvature modes}
\begin{figure}[tb]
  \includegraphics{fig/flatdiscontinuous.pdf}
  \caption{\label{fig:flatdiscontinuous}
    (a) Energy profile of the BITSS pathway on a 2D potential with flat regions.
        Blue and red dots denote the minima and transition state, respectively.
        Points of interest are labelled by i--iii (see text).
        Top inset: A zoomed in view around the transition state.
        Bottom inset: The pathway taken, with the edges of the flat regions marked by dashed lines.
    (b) Disconnectivity graph of the energy landscape for a seven-particle cluster with a discontinuous hard-core pair potential.
    (c) Plot of the discontinuous potential.
        The repulsive section of the standard Lennard-Jones potential is also shown in grey.
  }
\end{figure}

@@

\begin{thebibliography}{42}%

@@

@@

@@

@@

  {pages} {1} (\bibinfo {year} {2018})}\BibitemShut {NoStop}%

>>>>

% Figures
\usepackage{graphicx}
\graphicspath{{fig/}}

@@

More recently, it is increasingly being recognised that transition states are useful in mesoscale or macroscale systems.
Here, the minimum energy barriers provide important lower bounds to the energy input required for transitions to occur.
This has been used to understand failure in structural engineering applications \cite{Panter2019,Hutchinson2018}, for the development of super liquid-repellent surfaces \cite{Zhang2014,Panter2019b,Amabili2017}, and investigating locomotion through complex terrain for robotics \cite{Othayoth2020}.
Moreover, it is becoming desirable to tailor elastic deformation transitions to enable technologies such as advanced deployable structures \cite{Filipov2015,Zhai2018}, mechanical sensors and actuators \cite{Bertoldi2017,Truby2016,Chi2022,Bonfanti2020}, and energy absorbers \cite{Shan2015,Giri2021} to name but a few.

Existing transition state search methods generally fall into two categories, single- and double-ended methods.
Single-ended methods are initialised at a single state and attempt to climb to a nearby saddle point.
Examples include eigenvector following \cite{Cerjan1981}, the dimer method \cite{Heyden2005,Kastner2008,Zhang2016}, and climbing image methods \cite{E2007,Ren2013}.
Double-ended methods, such as the string method \cite{E2002,E2007} and doubly-nudged elastic band (DNEB) \cite{Trygubenko2004}, involve initialising a chain of states between two minima and then minimising the total energy of the chain.
The initial chain must use an appropriate interpolation, which can sometimes be complicated to produce \cite{Wales2012a}.
In addition to approximating the transition state, these double-ended methods also provide an estimate for the full transition pathway.

However, a large range of landscapes prove challenging or impossible to explore via these methods.
One key problem arises from the push towards larger and more complex systems \cite{Trefethen2013,Shalf2020,Alexander2020}, resulting in the need to develop algorithms that are more memory and computationally efficient, and can incorporate advanced optimisation strategies such as on-the-fly adaptive remeshing and coarse-graining.
Current double-ended methods require a chain of tens of states to gain an accurate approximation, and have issues with remeshing due to the different number of degrees of freedom for each state.
Single-ended methods are not well suited for identifying specific pathways and can spend a large amount of time searching for undesired transition states.
Another major challenge in studying complex energy landscapes relates to the presence of locally flat or discontinous regions, such as when considering patchy \cite{McMullen2018,Nguemaha2018,Chen2018b} and hard-body \cite{Richard2018,Santra2018} interactions in atomistic simulations, systems of polymer chains \cite{Mokkonen2016}, or collision constraints for macroscopic objects \cite{Wriggers2006}.
Flat zero-modes in the landscape pose issues for single-ended search methods that rely only upon local information.
Specialist treatment can sometimes be used such as in the case of global rotation and translation \cite{Page1988}, but they are thwarted by local zero-modes.
Meanwhile, current methods cannot typically be applied in the case of discontinuous potentials, or where the gradient is prohibitively expensive to compute, because they require continuous, differentiable optimisation functions.

\begin{figure}[htb]
  \includegraphics{toy2d.pdf}
  \caption{\label{fig:toy2d}
    (a) The trajectories of the two states under the BITSS method (orange line) for a simple 2D potential with two minima (blue dots).
        The transition state is shown by a red dot, and the minimum energy pathway by the dashed line.
    (b) A snapshot of the BITSS minimisation showing the driving forces on each state due to the energy constraint, $\bm{F}_E$, and distance constraint, $\bm{F}_D$, with $E_1 < E_2$ and $\mathrm{d}(\bm{x}_1,\bm{x}_2) < d_i$.
    (c) The final configuration of the BITSS method showing the two states in orange, the transition state in red, and the negative curvature eigenvector, $\bm{\hat{\tau}}$.
  }
\end{figure}

In this work we introduce a new method, the Binary Image Transition State Search (BITSS).
This is a double-ended method that uses two states that converge toward a transition state from either side, therefore combining aspects of the two traditional transition state search categories.
This has the advantage of not requiring an initial interpolation, unlike other double-ended methods.
Using a range of different applications, we will demonstrate how BITSS addresses each of the above challenges.

@@

To ensure that the transition state is located successfully, the constraint strengths $\kappa_d$ and $\kappa_e$ are updated as the algorithm proceeds using information from the system.
They are initially calculated at the start of each minimisation, and regularly recalculated throughout (once per 100 iterations is used in this work).
They are set using the following equations such that each term in the gradient of \cref{eq:bitss} is of a similar size,
\begin{gather}
  \kappa_e = \frac {\alpha} {2 E_B},
  \label{eq:ke}
  \\
  \kappa_d = \max \left(
    \frac {\sqrt{|\bm{\nabla} E_1|^2 + |\bm{\nabla} E_2|^2}} {2\sqrt{2} \beta d_i} \; , \;
    \frac{E_B}{\beta d_i^2} \right),
  \label{eq:kd}
\end{gather}
where $\bm{\nabla} E_1$ and $\bm{\nabla} E_2$ are the gradients of the energies of the two states, and $\alpha$ and $\beta$ are parameters with recommended values of $\alpha = 10$ and $\beta = 0.1$ (see supplementary information \footnote{See Supplemental Material at [URL] for derivations of the constraint coefficients, extra details for the example systems, results for the speed test against DNEB, and a comparison of the discontinuous energy landscape with the standard Lennard-Jones cluster}).
Here, $E_B$ is an estimation for the current energy barrier; evaluated using the difference between the highest energy along a linear interpolation between the two states and the average energy of the two states.

@@

%\section{Speed Comparison}
\begin{figure*}[tb]
  \includegraphics{speedtest.pdf}
  \caption{\label{fig:speedtest}
    The convergence to the transition state as a function of the number of gradient calculations using the BITSS (black line) and string methods for (a) a Lennard-Jones seven-particle cluster, (b) cylindrical shell buckling, and (c) wetting of a chemically-striped surface.
    The string method is repeated with a differing number of images along the string, the number of which is listed in the legend in (c).
    The shown configurations correspond to the two minimum energy states and the transition state, marked by an asterisk.
  }
\end{figure*}
As a result of the BITSS method using only two states it has the potential to be significantly faster than chain-of-states approaches which typically use tens of states.
This is tested here by measuring the speed of convergence to the transition state for BITSS in comparison with the string method and DNEB for several representative examples.

@@

%\section{Adaptive mesh}
\begin{figure}[tb]
  \includegraphics{adaptivemesh.pdf}
  \caption{\label{fig:adaptivemesh}
    (a) Snapshots of the BITSS method applied to the buckling of a cylinder with a changing mesh.
        The radial displacement relative the unbuckled cylinder is shown, as well as the underlying triangular mesh.
    (b) BITSS applied to the striped wetting example with different resolutions for the two states.
        Each grid cell denotes 50x50 lattice nodes.
        The zoomed axes shows the difference in the fluid interface between the two final states, as well as the approximated transition state (solid black line).
        This is compared to the transition state found using a high resolution (dashed line).
  }
\end{figure}

@@

%\section{Zero-curvature modes}
\begin{figure}[tb]
  \includegraphics{flatdiscontinuous.pdf}
  \caption{\label{fig:flatdiscontinuous}
    (a) Energy profile of the BITSS pathway on a 2D potential with flat regions.
        Blue and red dots denote the minima and transition state, respectively.
        Points of interest are labelled by i--iii (see text).
        Top inset: A zoomed in view around the transition state.
        Bottom inset: The pathway taken, with the edges of the flat regions marked by dashed lines.
    (b) Disconnectivity graph of the energy landscape for a seven-particle cluster with a discontinuous hard-core pair potential.
    (c) Plot of the discontinuous potential.
        The repulsive section of the standard Lennard-Jones potential is also shown in grey.
  }
\end{figure}

@@

\begin{thebibliography}{46}%

@@

\bibitem [{\citenamefont {Amabili}\ \emph {et~al.}(2017)\citenamefont
  {Amabili}, \citenamefont {Giacomello}, \citenamefont {Meloni},\ and\
  \citenamefont {Casciola}}]{Amabili2017}%
  {Amabili}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Giacomello}},
  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Meloni}},\ and\ \bibinfo
  {author} {\bibfnamefont {C.~M.}\ \bibnamefont {Casciola}},\ }\href
  {https://doi.org/10.1103/PhysRevFluids.2.034202} {\bibfield  {journal}
  {\bibinfo  {journal} {Physical Review Fluids}\ }\textbf {\bibinfo {volume}
  {2}},\ \bibinfo {pages} {034202} (\bibinfo {year} {2017})}\BibitemShut

@@

\bibitem [{\citenamefont {Zhang}\ \emph {et~al.}(2016)\citenamefont {Zhang},
  \citenamefont {Du},\ and\ \citenamefont {Zheng}}]{Zhang2016}%
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
  {Zhang}}, \bibinfo {author} {\bibfnamefont {Q.}~\bibnamefont {Du}},\ and\
  \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Zheng}},\ }\href
  {https://doi.org/10.1137/140972676} {\bibfield  {journal} {\bibinfo
  {journal} {SIAM Journal on Scientific Computing}\ }\textbf {\bibinfo {volume}
  {38}},\ \bibinfo {pages} {A528} (\bibinfo {year} {2016})}\BibitemShut

@@

\bibitem [{\citenamefont {Wales}\ and\ \citenamefont
  {Carr}(2012)}]{Wales2012a}%
  {Wales}}\ and\ \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
  {Carr}},\ }\href {https://doi.org/10.1021/ct3004832} {\bibfield  {journal}
  {\bibinfo  {journal} {Journal of Chemical Theory and Computation}\ }\textbf
  {\bibinfo {volume} {8}},\ \bibinfo {pages} {5020} (\bibinfo {year}

@@

  {pages} {032606} (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {M{\"{o}}kk{\"{o}}nen}\ \emph
  {et~al.}(2016)\citenamefont {M{\"{o}}kk{\"{o}}nen}, \citenamefont
  {Ala-Nissila},\ and\ \citenamefont {J{\'{o}}nsson}}]{Mokkonen2016}%
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {M{\"{o}}kk{\"{o}}nen}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
  {Ala-Nissila}},\ and\ \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {J{\'{o}}nsson}},\ }\href {https://doi.org/10.1063/1.4962167} {\bibfield
  {journal} {\bibinfo  {journal} {The Journal of Chemical Physics}\ }\textbf
  {\bibinfo {volume} {145}},\ \bibinfo {pages} {094901} (\bibinfo {year}
  {2016})}\BibitemShut {NoStop}%
